---
title: "Overview of Forecasting Accuracy"
author: "Dries Reynders"
format:
  html: 
    toc: true
    toc-depth: 4
    embed-resources: true
    footnotes-hover: true
    reference-location: margin
    self-contained: true
execute: 
  warning: false
  message: false
date: 2023-02-15
csl: sbr.csl
editor_options: 
  chunk_output_type: console
---

# Intro

This is an example Quarto document for html-output. Of course this could be rendered to pdf as well

## Subsection 1 {#sec-subs1}

Random text


### Subsubsection 1

More random text

### Subsubsection 2

Even more random text

## Subsection 2

Here, I would like to refer to @sec-subs1

# Main section

Let's get some data

```{r}
#| echo: true

data("mtcars")


```

The following chunk loads some packages (includes set to false, because loading packages often comes with a lot of unwanted output:

```{{r}}
#| include: false

library(tidyverse)

```

```{r}
#| include: false

library(tidyverse)

```

## A graph

```{r}
#| label: fig-first
#| out-width: 50%
#| fig-align: center
#| fig-cap: First graph


ggplot(mtcars, aes(hp, mpg ))+
  geom_point()+
  theme_bw()

```

